\begin{document}

\maketitle
\begin{frame}{Outline}
\setcounter{tocdepth}{1}
\tableofcontents
\end{frame}

\section{Review}
\label{sec-1}
\begin{frame}[label=sec-1-1]{Some basic themes}
\begin{enumerate}
\item Humanity and Divinity of Christ
\item Reason and revelation
\item Works and Grace
\item Spirit and Structure
\item Church and State
\end{enumerate}
\end{frame}
\begin{frame}[label=sec-1-2]{Nominalism vs. Realism (philosophical concepts)}
\begin{itemize}
\item the really real vs. knowing the individual rather than the universal
\item cf. John Duns Scotus
\item William of Ockham: showing reason's limitations
\item Ockham's ``razor''
\end{itemize}
\end{frame}

\begin{frame}[label=sec-1-3]{3 ``proofs'' for the existence of God}
\begin{itemize}
\item Anselm of Canterbury (11th c.)
\item Thomas Aquinas (13th c.)
\item B. Pascal (17th c.)
\end{itemize}
\end{frame}

\begin{frame}[label=sec-1-4]{14th c.}
\begin{itemize}[<+->]
\item The Avignon papacy transfers the seat of the Popes from Italy to France
\item The Great Famine of 1315-1317 kills millions of people in Europe.
\item Beginning of the Renaissance in Italy
\item The Hundred Years' War begins when Edward III of England lays claim to the French throne. (1337) The French recruit troops and ships in Genoa, Monaco, and Nice (1345–1346).
\item Black Death kills around a third of the population of Europe. (1347–1351).
\item The Great Schism of the West begins in 1378, eventually leading to 3 simultaneous popes.
\item The poet Petrarch coins the term \emph{Dark Ages} to describe the preceding 900 years in Europe, beginning with the fall of the Western Roman Empire in 476 through to the renewal embodied in the \alert{Renaissance}.
\end{itemize}
\end{frame}
\section{Reformation}
\label{sec-2}
\begin{frame}[label=sec-2-1]{Vocabulary}
\begin{itemize}[<+->]
\item nepotism: handing out positions to family members
\item pluralism: illegally hold more than one clergy position
\item ignorance: esp. of clergy
\item indulgences
\item absenteeism: clergy did not live in their diocese or parishes
\item worldliness: clergy got drunk, lead armies, fathered children, gambled etc.
\end{itemize}
\end{frame}

\begin{frame}[label=sec-2-2]{Vocabulary 2}
\begin{itemize}
\item humanism
\begin{itemize}
\item ?
\item ?
\item ?
\end{itemize}
\end{itemize}
\end{frame}
\begin{frame}[label=sec-2-3]{Luther, father of Reformation}
Who was this person who stands so huge at the beginning of our period?

\url{https://prezi.com/yfmiihckhjj0/martin-luther-reformation/}
\end{frame}
\begin{frame}[label=sec-2-4]{Read for Monday}
Excerpts from Luther's ``Freedom of a Christian'' (found on the class web site)
\begin{itemize}
\item \url{https://sites.google.com/site/relg317f15/}

\item \url{http://richard-hooker.com/sites/worldcultures/REFORM/FREEDOM.HTM}
\end{itemize}
\end{frame}
% Emacs 24.3.1 (Org mode 8.2.3c)
\end{document}
